Automated Verification of Stand-alone Solar Photovoltaic Systems: Optimal Sizing and Project Validation

### 2020-02-02 Tex/Old text/abstract.tex
thesis update

\thispagestyle{plain}
\begin{center}
%    \Large
%    \textbf{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems}
%    
%    \vspace{0.4cm}
%    \large
%    Optimal Sizing and Project Validation
%    
%    \vspace{0.4cm}
%    \textbf{Alessandro Trindade}
%    
    \vspace{0.9cm}
    \textbf{Abstract}
\end{center}
With decreasing costs and increasing performance, the deployment of renewable energy systems is growing faster now than in the past decade. In 2017, for the first time, the number of people without access to electricity dipped below 1 billion, but trends in energy access still fall short of global goals. Particular attention is given to stand-alone solar photovoltaic systems in rural areas or where grid extension is unfeasible. Tools to evaluate or to size electrification projects are available, but they are based on simulations that do not cover all aspects of the design space. However, the use of formal methods to model and validate any system has grown with time, mainly to find bugs in sophisticated hardware and software systems: they aim to establish system correctness with mathematical rigor. The use of formal methods in electrical systems is a new subject, with published research spanning only the last four years. Moreover, the use of automated synthesis in order to obtain optimal sizing of solar photovoltaic systems has never been done before. This thesis marks the achievement of two major goals: first, the application of software model checking to verify formally the design of a stand-alone solar photovoltaic system, including solar panel, charge controller, battery, inverter, and electric load; second, a sound, automated approach to obtaining optimal sizing of stand-alone photovoltaic systems using program synthesis. For the formal verification, we used case studies from real photovoltaic systems deployed in five different sites, ranging from 700 W to 1.200 W, in order to evaluate the proposed approach and to compare  it with a specialized simulation tool. Different verification tools are evaluated also, in order to compare performance and soundness. Data from practical applications show the effectiveness of our proposed approach, where specific conditions that lead to failures in a photovoltaic solar system are detailed only by the automated verification method. In addition, for the use of program synthesis, we propose a variant of the counterexample guided inductive synthesis (CEGIS) approach. This variant has two phases linking the technical and the cost analysis. First, we synthesize a feasible candidate based on power reliability, but which may not attain the lowest cost. Second, the candidate is then verified iteratively with a lower bound cost via symbolic model checking. If the verification step succeeds, the lower bound is adjusted; if it fails, a counterexample provides the optimal solution. The proposed synthesis method is novel and unprecedented as it streamlines the design of photovoltaic systems. Experimental results using seven case studies demonstrate that our synthesis method can produce optimal system sizing within an acceptable run-time. We also present a comparison with a specialized simulation tool over real photovoltaic systems in order to show the effectiveness of our approach, which can provide a more detailed and accurate solution than the simulation tool.

>>>>

\thispagestyle{plain}
\begin{center}
%    \Large
%    \textbf{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems}
%    
%    \vspace{0.4cm}
%    \large
%    Optimal Sizing and Project Validation
%    
%    \vspace{0.4cm}
%    \textbf{Alessandro Trindade}
%    
    \vspace{0.9cm}
    \textbf{Abstract}
\end{center}
With decreasing costs and increasing performance, the deployment of renewable energy systems is now growing faster than in the past decade. In 2017, for the first time, the number of people without access to electricity dipped below 1 billion, but trends in energy access still fall short of global goals. Particular attention is given to stand-alone solar photovoltaic systems in rural areas or where grid extension is unfeasible. Tools to evaluate or to size electrification projects are available, but they are based on simulations that do not cover all aspects of the design space. However, the use of formal methods to model and validate any system has grown with time, mainly to find bugs in sophisticated hardware and software systems: they aim to establish system correctness with mathematical rigor. The use of formal methods in electrical systems is a new subject, with published research spanning only the last four years. Moreover, the use of automated synthesis in order to obtain optimal sizing of solar photovoltaic systems has never been done before. This thesis marks the achievement of two major goals: first, the application of software model checking to verify formally the design of a stand-alone solar photovoltaic system, including solar panel, charge controller, battery, inverter, and electric load; second, a sound, automated approach to obtaining optimal sizing of stand-alone photovoltaic systems using program synthesis. For the formal verification, we used case studies from real photovoltaic systems deployed in five different sites, ranging from 700 W to 1.200 W, in order to evaluate the proposed approach and to compare  it with a specialized simulation tool. Different verification tools are evaluated also, in order to compare performance and soundness. Data from practical applications show the effectiveness of our proposed approach, where specific conditions that lead to failures in a photovoltaic solar system are detailed only by the automated verification method. In addition, for the use of program synthesis, we propose a variant of the counterexample guided inductive synthesis (CEGIS) approach. This variant has two phases linking the technical and the cost analysis. First, we synthesize a feasible candidate based on power reliability, but which may not attain the lowest cost. Second, the candidate is then verified iteratively with a lower bound cost via symbolic model checking. If the verification step succeeds, the lower bound is adjusted; if it fails, a counterexample provides the optimal solution. The proposed synthesis method is novel and unprecedented as it streamlines the design of photovoltaic systems. Experimental results using seven case studies demonstrate that our synthesis method can produce optimal system sizing within an acceptable run-time. We also present a comparison with a specialized simulation tool over real photovoltaic systems in order to show the effectiveness of our approach, which can provide a more detailed and accurate solution than the simulation tool.

### 2020-02-02 Tex/Old text/chapters/chapter02.tex
thesis update

\begin{itemize}
\item In modeling, a model (normally mathematical) of the system is created; 
\item In specification, normally a list of properties to be satisfied by the system is established, i.e., the requirements, such as reliability to performance, for example; 
\item Normally it is expressed in temporal logic form ($CTL$); 
\item The model checking is the verification method itself. 
\end{itemize}

@@

\begin{itemize}
\item Stand-alone systems, where the energy is generated and consumed in the same place and which does not interact with the main grid. Normally, the electricity consuming/utilizing device is part of the system, i.e., solar home systems, solar street lighting systems, solar lanterns, and solar power plants; 
\item Grid-connected systems, where the solar PV system is connected to the grid. The grid-connected system can be either a grid-tied system, which can only feed power into the grid, with the result that this system cannot deliver power locally during blackouts and emergencies since these systems have to be completely disconnected from the grid and have to be shut down as per national and international electrical safety standards. Some grid-connected PV systems, with energy storage, can also provide power locally in an islanding mode; 
\item Solar PV hybrid systems: In a hybrid system, another source(s) of energy, such as wind, biomass or diesel, can work together with the solar PV system to provide the required demand. In this type of system, the main objective is to bring more reliability into the overall system at an affordable cost by adding one or more energy sources.
\end{itemize}

@@

According to \cite{Pinho}, PV systems that are used to feed loads with low variation of consumption can be sized to operate without the controller. This is known as a self-regulated stand-alone PV system with battery. However, PV panel voltage must be compatible with the battery voltage. Normally, the bank of batteries is oversized in relation to the PV panel and to the load. The drawback is the operation of the batteries, normally overloaded or with excessive discharges (that can damage the batteries).

@@

The performance of photovoltaic systems (solar cell/panels), that is, the output current/voltage curve ($I-V$ curve), is usually studied using an equivalent circuit model. This equivalent circuit consists of a current source with one or two diodes connected in parallel, and up to two resistors, one connected in parallel and the other one in series, to take into account energy losses in this model \cite{Cubas}. Based on these electronic components, four basic configurations are normally used when studying photovoltaic systems, as shown in Fig. \ref{fig:equivckt}.

@@

Several models have been presented in the literature. However, regardless of the model, normally the following parameters are normally required\cite{Mellit}:

@@

\subsubsection{The inverter model}
As shown by \cite{Mellit}, PV arrays produce DC and therefore when the PV system contains an AC load, DC/AC conversion is required. An inverter is a converter, where the power flows from the DC to the AC side, having DC voltage as input and producing AC voltage as output. The role of the inverter is to keep the voltage constant on the AC side, i.e., at the rated voltage (127 V or 220 V, for example), and to convert input power $ P_{in} $ into  output power $ P_{out} $ with the best possible efficiency \cite{Hansen}.

@@

\noindent where the variable $autonomy$ is a design definition and normally has a value ranging from $6$ to $48$ hours; 
$ V_{system} $ is the DC voltage of the bus, and $ DOD $ is the battery depth of discharge (considered the maximum of 25\% here).
%
Second, the total number of batteries is computed as

>>>>

\begin{itemize}
\item In modeling, a model (usually mathematical) of the system is created; 
\item In specification, usually a list of properties to be satisfied by the system is established, i.e., the requirements, such as reliability to performance, for example; 
\item It is expressed usually in temporal logic form ($CTL$); 
\item The model checking is the verification method itself. 
\end{itemize}

@@

\begin{itemize}
\item Stand-alone systems, where the energy is generated and consumed in the same place and which does not interact with the main grid. Usually, the electricity consuming/utilizing device is part of the system, i.e., solar home systems, solar street lighting systems, solar lanterns, and solar power plants; 
\item Grid-connected systems, where the solar PV system is connected to the grid. The grid-connected system can be either a grid-tied system, which can only feed power into the grid, with the result that this system cannot deliver power locally during blackouts and emergencies since these systems have to be completely disconnected from the grid and have to be shut down as per national and international electrical safety standards. Some grid-connected PV systems, with energy storage, can also provide power locally in an islanding mode; 
\item Solar PV hybrid systems: In a hybrid system, another source(s) of energy, such as wind, biomass or diesel, can work together with the solar PV system to provide the required demand. In this type of system, the main objective is to bring more reliability into the overall system at an affordable cost by adding one or more energy sources.
\end{itemize}

@@

According to \cite{Pinho}, PV systems that are used to feed loads with low variation of consumption can be sized to operate without the controller. This is known as a self-regulated stand-alone PV system with battery. However, PV panel voltage must be compatible with the battery voltage. Usually, the bank of batteries is oversized in relation to the PV panel and to the load. The drawback is the operation of the batteries, usually overloaded or with excessive discharges (that can damage the batteries).

@@

The performance of photovoltaic systems (solar cell/panels), that is, the output current/voltage curve ($I-V$ curve), is usually studied using an equivalent circuit model. This equivalent circuit consists of a current source with one or two diodes connected in parallel, and up to two resistors, one connected in parallel and the other one in series, to take into account energy losses in this model \cite{Cubas}. Based on these electronic components, four basic configurations are usually used when studying PV systems, as shown in Fig. \ref{fig:equivckt}.

@@

Several models have been presented in the literature. However, regardless of the model, the following parameters are usually required\cite{Mellit}:

@@

\subsubsection{The Inverter Model}
As shown by \cite{Mellit}, PV arrays produce DC and therefore when the PV system contains an AC load, DC/AC conversion is required. An inverter is a converter, where the power flows from the DC to the AC side, having DC voltage as input and producing AC voltage as output. The role of the inverter is to keep the voltage constant on the AC side, i.e., at the rated voltage (127 V or 220 V, for example), and to convert input power $ P_{in} $ into  output power $ P_{out} $ with the best possible efficiency \cite{Hansen}.

@@

\noindent where the variable $autonomy$ is a design definition and usually has a value ranging from $6$ to $48$ hours; 
$ V_{system} $ is the DC voltage of the bus, and $ DOD $ is the battery depth of discharge (considered the maximum of 25\% here).
%
Second, the total number of batteries is computed as

### 2020-02-02 Tex/Old text/chapters/introduction.tex
thesis update

Among the many CPS systems, in this thesis we will focus specifically on energy generation by stand-alone solar PV systems. Lack of access to clean and affordable energy is considered a core dimension of poverty~\cite{Hussein2012}. Nevertheless, the world is progressing; in particular, the number of people without electricity access fell below the 1 billion threshold for the first time in 2017~\cite{IEAweo2018}. The share of people without access to electricity from Africa is 58\%, while 19\% of the share comes from developing Asia, and 31\% from Latin America~\cite{IEAweo2018}. Numbers from Brazil show the aim to electrify 270 isolated areas and 2.7 million people by 2023~\cite{EPE2018}. 
Furthermore, there is a close relationship between the lack of energy and the low HDI (Human Development Index) of those localities~\cite{Coelho}. It follows that, increased access to energy allows economic growth and poverty alleviation \cite{Karekesi}.

The use of the power generation technology with a renewable energy source is developing rapidly due to the industrial development~\cite{Yatimi}. Renewable energy leads to advances all over the world by protecting the environment: it is clean (low greenhouse gas emissions), operates silently, long-lasting, low maintenance costs, zero fuel costs and an  inexhaustible supply~\cite{Noroozian}. Renewable energy, and particularly power generated from solar energy using photovoltaic (PV) panels, has emerged as an alternative to fossil or nuclear fuel generation.

Renewable sources of energy include hydro, wind, and solar PV. According to~\cite{SEIA} and~\cite{Chauhan}, solar is the most abundant source of renewable energy on earth. In solar PV systems, solar radiation is captured from the sun and turned into electricity using solar PV cells made of silicon and other materials.

Nothing more than a niche market only a few years ago, solar PV systems have become a mainstream electricity provider, with an  approximate 50\% increase in new PV installations from 2016 to 2017~\cite{EPIA}, although this growth is not driven by stand-alone systems.

@@

%-------------------------------------------------
\section{The Problem}
%-------------------------------------------------

@@

The main objective of this thesis is thus to prove that is possible to use formal verification method to validate the design and to use formal synthesis method to optimize the sizing of stand-alone solar PV systems. Secondarily, we aim to develop two tools written in ANSI-C and platform-independent, based on the the new proposed techniques in order to evaluate different state-of-the art symbolic software verifiers (and solvers), and comparing the experimental results with simulation tool and with data from real PV systems deployed in riverside communities. Always evaluating correctness and performance.

@@

%This Ph.D. thesis has the novelty of create an automated verification approach to solar PV systems. The research led to the creation of two software to carry out computational experiments and to compare with commercial tools and real data from systems installed in the field. One software to perform automated verification and validate the solar PV systems behavior, and another one to produce the optimal sizing of solar PV systems through automated synthesis program.
In this study, a mathematical model of each component of a stand-alone PV system: the solar panel, charge controller, batteries, inverter, and electrical load, is used for formal verification. The intended behavior of each system component can be verified and observed with the support of the formal models, as a joint operation of the components, which in this case represents the operation of the solar PV system itself. The project requirements, such as battery autonomy and power demand, as well as the weather conditions, such as solar irradiance and ambient temperature, are provided for the proposed tool and automatically verified during the formal process. The model checking tool reports in which conditions a system does not meet the user requirements. A key benefit to this approach is that it helps in the detection of flaws in the design phase of system development, thereby considerably improving system reliability~\cite{Akram2018}. 
%Those programs are viewed as mathematical objects, and the use of 
%Bringing this technique to solar PV systems, i
%Instead of finding a bug or a software inconsistence, t
%Here it will be verified a stand-alone PV system, with all their components modeled mathematically,   
%Among the non-deterministic variables considerated, is it worth to mention environmental issues (like solar irradiance, and temperature), and the variation of the electrical load (power consumption) or the battery' state of charge, as shown in Fig.\ref{fig:flowchartgeneral}.

The proposed method is applied by means of an algorithm implemented using the C programming language. We performed computing experimentation in three state-of-the-art model checkers to formally verify PV designs, aiming to evaluate performance and correctness: the C Bounded Model Checker (CBMC)~\cite{Kroening}, the Efficient SMT-based Bounded Model Checker (ESBMC)~\cite{esbmc2018}, and the Configurable Program Analysis Checker (CPAchecker).
%three the efficient SMT-based bounded model checker (ESBMC)~\cite{esbmc2018}. %, which allows one to incrementally verify a PV system as an imperative program using a fragment of decidable first-order theories~\cite{DBLP:books/daglib/0019162}.

With respect to the \textbf{automated synthesis} method and optimal sizing of stand-alone PV systems, this thesis presents the development of a variant of counterexample guided inductive synthesis (CEGIS) that uses commercial equipment data, user requirements and weather conditions as input. Given a correctness specification $\sigma$, the proposed method uses $\sigma$ as a starting point and then iteratively produces a sequence of candidate solutions that satisfy $\sigma$ with respect to power reliability. In particular, in each iteration, we synthesize the sizing of stand-alone PV systems, which may not attain the lowest cost. The candidate solution is then verified via symbolic model checking with a lower bound that serves as the minimum cost of reference; if the verification step does not fail, the lower bound is adjusted. If it fails, then a counterexample is provided with an optimal sizing that meets power reliability and system cost requirements. Note that in this study, the focus is not on new criteria or even optimization objectives. Instead, the novelty lies in the practical approach to the pursuit of the optimal solution of PV systems using formal methods, which outperforms existing state-of-the-art simulation tools.

@@

Our approach deals with the theoretical and pragmatic aspects of using model checking in stand-alone solar PV systems in order to tackle two issues: validation of a designed system and optimal sizing. Validation when a sized system must be evaluated in order to decide if it meets the design requirements and user needs; and optimization when there are just the requirements and the PV system is not sized, or when the user tries to obtain a better solution that he or she has. We thus develop algorithms and the corresponding tools; we then evaluate them through several case studies, in a comparison with specialized simulation tools, supported by real data from deployed PV systems.

@@

\begin{figure}[h]
\includegraphics[width=0.85\textwidth]{verification_outline2.png}
\centering
\caption{Proposed validation of PV systems via automated verification.}
\label{fig:validation_outline} 
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.85\textwidth]{optimization_outline2.png}
\centering
\caption{Proposed optimal sizing of PV systems via automated synthesis.}
\label{fig:optimization_outline} 
\end{figure}

@@

With respect to the proposed \textbf{automated verification} method, this thesis makes three main contributions: \textbf{first}, we propose an algorithm written in the C programming language which implements the automated verification method to formally check the sizing and the operation of a given stand-alone PV system; \textbf{second}, we evaluate the verification method by comparing three state-of-the-art model checkers in five real case studies; \textbf{third}, experimental results show that this approach can find subtle design errors in stand-alone PV systems not easily detected by other approaches based on simulation.

@@

It is worth mentioning that we decided, with the consent of the coordination of the graduate program, that this document would be written in the active voice and entirely in English. The choice of the active voice comes from the choice of the English language, which in the thesis studied was always presented in the active voice. The decision to use the English language stems from the possibility that the research may lead to further developments and activities, without the need for translation.

>>>>

Among the many CPS systems, in this thesis we will focus specifically on energy generation by stand-alone solar PV systems. The lack of access to clean and affordable energy is considered a core dimension of poverty~\cite{Hussein2012}. Nevertheless, the world is progressing; in particular, the number of people without electricity access fell below the 1 billion threshold for the first time in 2017~\cite{IEAweo2018}. The share of people without access to electricity from Africa is 58\%, while 19\% of the share comes from developing Asia, and 31\% from Latin America~\cite{IEAweo2018}. Numbers from Brazil show the aim to electrify 270 isolated areas and 2.7 million people by 2023~\cite{EPE2018}. 
Furthermore, there is a close relationship between the lack of energy and the low HDI (Human Development Index) of those localities~\cite{Coelho}. It follows that, increased access to energy allows economic growth and poverty alleviation \cite{Karekesi}.

The use of the power generation technology with renewable energy sources is developing rapidly due to the industrial development~\cite{Yatimi}. Renewable energy leads to advances all over the world by protecting the environment: it is clean (low greenhouse gas emissions), operates silently, long-lasting, low maintenance costs, zero fuel costs and an  inexhaustible supply~\cite{Noroozian}. Renewable energy, and particularly power generated from solar energy using photovoltaic (PV) panels, has emerged as an alternative to fossil or nuclear fuel generation.

Renewable sources of energy include hydro, wind, and solar PV. According to~\cite{SEIA} and~\cite{Chauhan}, the sun is the most abundant source of renewable energy on earth. In solar PV systems, solar radiation is captured from the sun and turned into electricity using solar PV cells made of silicon and other materials.

Nothing more than a niche market only a few years ago, solar PV systems have become a mainstream electricity provider, with an  approximate 50\% increase (or 100.9 GW)in new PV installations from 2016 to 2017~\cite{EPIA}, although this growth is not driven by stand-alone systems.

@@

%-------------------------------------------------
\section{The Problem}
%-------------------------------------------------
%
%Global demand for energy is predicted to increase in the coming decades. The International Energy Agency’s 2007 World Energy Outlook states that between now and 2030:Global energy needs are expected to grow, with fossil fuels remaining the dominant source.
%Between 2005 and 2030, energy needs are projected to expand by 55 per cent, with demand increasing from 11.4 billion tons of oil equivalent to 17.7 billion. Between 2005 and 2030, energy consumption is expected to increase by 50 per cent, with the bulk of the demand coming from developing countries. Oil, coal and gas together account for the majority of global primary energy consumption. However, 
%

Energy access is an important issue because it can put together economic growth, human development and environmental sustainability. Energy has long been recognized as essential for humanity to develop and thrive, but the adoption in 2015 by 193 countries of a goal to ensure access to affordable, reliable, sustainable and modern energy for all by 2030, as part of the new United Nations Sustainable Development Goals (SDGs), marked a new level of political recognition~\cite{IEAweo2018}.

The International Energy Agency (IEA) revealed that from 2000 to 2016 nearly all of those who gained access to electricity worldwide did so through new grid connections, mostly with power generation from fossil fuels. Over the last five years, however, renewable sources have started to gain ground, as have off-grid and mini-grid systems. By 2030, renewable energy sources power over 60\% of new access, and off-grid and mini-grid systems provide the means for almost half of new access~\cite{IEAweo2018}. Moreover, if we focus into the Amazonas State, Brazil, where around 5\% of the population in 2018 were isolated in 2,261 communities (or 41,167 houses), without electrical energy or even road access, using just boats by rivers, therefore energy is a very important issue.

Under this scenario, it is important to have very well designed PV systems that won't fail when installed in the field and at the lowest acquisition cost possible.

@@

The main objective of this thesis is thus to prove that is possible to use formal verification method to validate the design and to use formal synthesis method to optimize the sizing of stand-alone solar PV systems. Secondarily, we aim to develop two tools written in ANSI-C and platform-independent, based on the the new proposed techniques in order to evaluate different state-of-the art symbolic software verifiers (and solvers), and comparing the experimental results with simulation tool and with data from real PV systems installed in riverside communities. Always evaluating correctness and performance.

@@

%This Ph.D. thesis has the novelty of create an automated verification approach to solar PV systems. The research led to the creation of two software to carry out computational experiments and to compare with commercial tools and real data from systems installed in the field. One software to perform automated verification and validate the solar PV systems behavior, and another one to produce the optimal sizing of solar PV systems through automated synthesis program.
In this study, a mathematical model of each component of a stand-alone PV system: the solar panel, charge controller, batteries, inverter, and electrical load, is used for formal verification. The intended behavior of each system component can be verified and observed with the support of formal models, as a joint operation of the components, which in this case represents the operation of the solar PV system itself. The project requirements, such as battery autonomy and power demand, as well as the weather conditions, such as solar irradiance and ambient temperature, are provided for the proposed tool and automatically verified during the formal process. The model checking tool reports in which conditions a system does not meet the user requirements. A key benefit to this approach is that it helps in the detection of flaws in the design phase of system development, thereby considerably improving system reliability~\cite{Akram2018}. 
%Those programs are viewed as mathematical objects, and the use of 
%Bringing this technique to solar PV systems, i
%Instead of finding a bug or a software inconsistence, t
%Here it will be verified a stand-alone PV system, with all their components modeled mathematically,   
%Among the non-deterministic variables considerated, is it worth to mention environmental issues (like solar irradiance, and temperature), and the variation of the electrical load (power consumption) or the battery' state of charge, as shown in Fig.\ref{fig:flowchartgeneral}.

The proposed technique is applied by means of an algorithm implemented using the C programming language. We performed computing experimentation in three state-of-the-art model checkers to formally verify PV designs, aiming to evaluate performance and correctness: the C Bounded Model Checker (CBMC)~\cite{Kroening}, the Efficient SMT-based Bounded Model Checker (ESBMC)~\cite{esbmc2018}, and the Configurable Program Analysis Checker (CPAchecker).
%three the efficient SMT-based bounded model checker (ESBMC)~\cite{esbmc2018}. %, which allows one to incrementally verify a PV system as an imperative program using a fragment of decidable first-order theories~\cite{DBLP:books/daglib/0019162}.

With respect to the \textbf{automated synthesis} technique and optimal sizing of stand-alone PV systems, this thesis presents the development of a variant of counterexample guided inductive synthesis (CEGIS) that uses commercial equipment data, user requirements and weather conditions as input. Given a correctness specification $\sigma$, the proposed method uses $\sigma$ as a starting point and then iteratively produces a sequence of candidate solutions that satisfy $\sigma$ with respect to power reliability. In particular, in each iteration, we synthesize the sizing of stand-alone PV systems, which may not attain the lowest cost. The candidate solution is then verified via symbolic model checking with a lower bound that serves as the minimum cost of reference; if the verification step does not fail, the lower bound is adjusted. If it fails, then a counterexample is provided with an optimal sizing that meets power reliability and system cost requirements. Note that in this study, the focus is not on new criteria or even optimization objectives. Instead, the novelty lies in the practical approach to the pursuit of the optimal solution of PV systems using formal methods, which outperforms existing state-of-the-art simulation tools.

Simply put, we aim to present new techniques that will ensure that you have good designs of stand-alone solar PV systems, that do not cause a drop in energy supply, and that these systems are designed at the lowest possible cost.

@@

Our approach deals with the theoretical and pragmatic aspects of using model checking in stand-alone solar PV systems in order to tackle two issues: validation of a designed system and optimal sizing. Validation when a sized system must be evaluated in order to decide if it meets the design requirements and user needs; and optimization when there are just the requirements and the PV system is not sized, or when the user tries to obtain a better solution that he or she has. We thus develop algorithms and the corresponding tools; we then evaluate them through several case studies, in a comparison with specialized simulation tools, supported by real data from installed PV systems.

@@

\begin{figure}[h]
\includegraphics[width=0.85\textwidth]{verification_outline3.png}
\centering
\caption{Proposed validation of PV systems via automated verification.}
\label{fig:validation_outline} 
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.85\textwidth]{optimization_outline3.png}
\centering
\caption{Proposed optimal sizing of PV systems via automated synthesis.}
\label{fig:optimization_outline} 
\end{figure}

@@

With respect to the proposed \textbf{automated verification} technique, this thesis makes three main contributions: \textbf{first}, we propose an algorithm written in the C programming language which implements the automated verification method to formally check the sizing and the operation of a given stand-alone PV system; \textbf{second}, we evaluate the verification technique by comparing three state-of-the-art model checkers in five real case studies; \textbf{third}, experimental results show that this approach can find subtle design errors in stand-alone PV systems not easily detected by other approaches based on simulation.

@@

It is worth mentioning that we decided, with the consent of the coordination of the graduate program, that this document would be written in the active voice and entirely in English. The choice of the active voice comes from the choice of the English language, which in the thesis studied was always presented in the active voice. The decision to use the English language stems from the possibility that the research may lead to further developments and activities,  in the possibility of reaching a larger number of people and helping in the internationalization of the Federal University of Amazonas.

### 2020-02-02 Tex/Old text/resumo.tex
thesis update

Com custos decrescentes e com melhoria de desempenho, a implantação de sistemas de energia renovável está crescendo cada vez mais rapidamente no mundo. Em 2017, pela primeira vez, o número de pessoas sem acesso a eletricidade ficou abaixo de 1 bilhão, mas os dados quanto à universalização do acesso a energia ficaram aquém das metas globais. Particular atenção é dada aos sistemas isolados solares fotovoltaicos em áreas rurais ou onde a extensão da rede é inviável. Ferramentas para avaliar ou dimensionar projetos de eletrificação estão disponíveis, mas elas são baseadas em simulações que não cobrem todos os aspectos do espaço de projeto. Por outro lado, o uso de métodos formais para modelar e validar qualquer tipo de sistema está crescendo com o tempo, principalmente para encontrar bugs em sistemas complexos de \textit{hardware} e \textit{software}: seu objetivo é estabelecer a corretude do sistema com rigor matemático. O uso de métodos formais em sistemas elétricos é um assunto recente, com pesquisas sendo publicadas apenas nos últimos quatro anos. Além disso, a síntese automatizada nunca foi usada antes para obter um ótimo dimensionamento de sistemas solares fotovoltaicos . Esta tese marca duas conquistas principais: (1) a primeira aplicação de verificação de modelos de \textit{software} para verificar o projeto de um sistema isolado solar fotovoltaico, incluindo painel solar, controlador de carga, bateria, inversor e carga elétrica; e (2) uma abordagem confiável e automatizada para obter o dimensionamento ótimo de sistemas fotovoltaicos usando a síntese de programas onde cada componente e função de um sistema solar fotovoltaico é descrito, incluindo suas propriedades, e o modelo comportamental que representa o dimensionamento ótimo é sintetizado automaticamente. Relacionado à verificação formal, estudos de caso de sistemas fotovoltaicos reais instalados em cinco localidades diferentes são usados para avaliar a abordagem proposta e para compará-la com ferramenta de simulação especializada. Diferentes ferramentas de verificação são avaliadas também, a fim de comparar o desempenho e a confiabilidade dos resultados. Dados de aplicações práticas mostram a eficácia da abordagem proposta, onde condições específicas que levam a falhas em um sistema solar fotovoltaico são detalhadas apenas pelo método de verificação automatizado. Além disso, em relação ao uso da síntese de programas, propõe-se uma variante do método de síntese indutiva guiada por contraexemplos (CEGIS), com duas fases bem definidas: primeiro, ele sintetiza o dimensionamento de sistemas fotovoltaicos baseados em confiabilidade de energia, mas que pode não alcançar o menor custo; segundo, a solução proposta é então verificada iterativamente com um limite inferior via verificação de modelo simbólico. Se a etapa de verificação não falhar, o limite inferior será ajustado; e se falhar, o contraexemplo é fornecido com o dimensionamento ótimo, vinculando assim a resposta técnica da primeira fase à análise de custo da segunda fase. Os dados de equipamentos comerciais de diferentes fabricantes são fornecidos ao mecanismo de síntese e as soluções candidatas são derivadas da análise financeira do dimensionamento obtido. O método de síntese proposto é novo e sem precedentes para simplificar o projeto de sistemas fotovoltaicos. Resultados experimentais usando sete estudos de caso mostram que o nosso método de síntese é capaz de produzir em um tempo de execução aceitável o dimensionamento ótimo do sistema fotovoltaico, e um comparativo com uma ferramenta de simulação especializada e sistemas fotovoltaicos reais mostra a eficácia da abordagem adotada.

>>>>

Com custos decrescentes e com melhoria de desempenho, a implantação de sistemas de energia renovável está crescendo cada vez mais rapidamente no mundo. Em 2017, pela primeira vez, o número de pessoas sem acesso a eletricidade ficou abaixo de 1 bilhão, mas os dados quanto à universalização do acesso a energia ficaram aquém das metas globais. Particular atenção é dada aos sistemas isolados solares fotovoltaicos em áreas rurais ou onde as elevadas extensões tornam a rede inviável. Ferramentas para avaliar ou dimensionar projetos de eletrificação estão disponíveis, mas elas são baseadas em simulações que não cobrem todos os aspectos do espaço de projeto. Por outro lado, o uso de métodos formais para modelar e validar qualquer tipo de sistema está crescendo com o tempo, principalmente para encontrar "bugs" em sistemas complexos de \textit{hardware} e \textit{software}: seu objetivo é estabelecer a corretude do sistema com rigor matemático. O uso de métodos formais em sistemas elétricos é um assunto recente, com pesquisas sendo publicadas apenas nos últimos quatro anos. Além disso, a síntese automatizada nunca foi usada antes para se obter um ótimo dimensionamento de sistemas solares fotovoltaicos. Esta tese marca duas conquistas principais: (1) a primeira aplicação de verificação de modelos de \textit{software} para verificar o projeto de um sistema isolado solar fotovoltaico, incluindo painel solar, controlador de carga, bateria, inversor e carga elétrica; e (2) uma abordagem confiável e automatizada para obter o dimensionamento ótimo de sistemas fotovoltaicos usando a síntese de programas onde cada componente e função de um sistema solar fotovoltaico é descrito, incluindo suas propriedades, e o modelo comportamental que representa o dimensionamento ótimo é sintetizado automaticamente. Relacionado à verificação formal, estudos de caso de sistemas fotovoltaicos reais instalados em cinco localidades diferentes são usados para avaliar a abordagem proposta e para compará-la com ferramenta de simulação especializada. Diferentes ferramentas de verificação são avaliadas também, a fim de comparar o desempenho e a confiabilidade dos resultados. Dados de aplicações práticas mostram a eficácia da abordagem proposta, onde condições específicas que levam a falhas em um sistema solar fotovoltaico são detalhadas apenas pelo método de verificação automatizado. Além disso, em relação ao uso da síntese de programas, propõe-se uma variante do método de síntese indutiva guiada por contraexemplos (CEGIS), com duas fases bem definidas: primeiro, ele sintetiza o dimensionamento de sistemas fotovoltaicos baseados em confiabilidade de energia, mas que pode não alcançar o menor custo; segundo, a solução proposta é então verificada iterativamente com um limite inferior via verificação de modelo simbólico. Se a etapa de verificação não falhar, o limite inferior será ajustado; e se falhar, o contraexemplo é fornecido com o dimensionamento ótimo, vinculando assim a resposta técnica da primeira fase à análise de custo da segunda fase. Os dados de equipamentos comerciais de diferentes fabricantes são fornecidos ao mecanismo de síntese e as soluções candidatas são derivadas da análise financeira do dimensionamento obtido. O método de síntese proposto é novo e sem precedentes para simplificar o projeto de sistemas fotovoltaicos. Resultados experimentais usando sete estudos de caso mostram que o nosso método de síntese é capaz de produzir em um tempo de execução aceitável o dimensionamento ótimo do sistema fotovoltaico, e um comparativo com uma ferramenta de simulação especializada e sistemas fotovoltaicos reais mostra a eficácia da abordagem adotada.
